Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-02 nge.tex
Address Shantenu's comments to 6.0/6.1

\jhanote{not sure MPI script is optimal description or ambigious.} Multiple
generations would also relax two assumptions of the current execution model:
knowing the number of simulations before submitting the MPI script, and having
a fixed number of events per simulation (100 at the moment). Pilots would
enable the scheduling of simulations independently from whether they were
available at the moment of submitting the pilot. Further, simulations with a
varying number of events could be scheduled on a pilot, depending on the
amount of remaining walltime and the distribution of execution time per event,
as shown in \S\ref{ssec:panda_titan}, Figure~\ref{fig:100event-distrib}.
These capabilities would increase the efficiency of the PanDA Broker when
there is a large difference between the number of cores and walltime.

@@

\jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
dynamically determined tasks that constitute workloads." This definition put
the focus on tasks and not workloads, which is a level above in the RADICAL
conceptual layer diagram.} NGE is a runtime system designed to execute
heterogeneous and dynamic workloads on diverse resources. Fig.~\ref{fig:arch-
overview} illustrates its current architecture as deployed on Titan: the two
management modules (Pilot and Unit) represent a simplified version of the PanDA Broker while
the agent module is the pilot submitted to Titan and executed on its worker
nodes. The communication between PanDA Broker and PanDA Server is abstracted
away as it is not immediately useful to evaluate the performance and
capabilities of a pilot on Titan.

@@

NGE exposes an API for the application  \jhanote{workload management?} layer
to describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager.   The PilotManager submits pilots to Titan's PBS
\jhanote{for consistency should we say Moab?} batch system via SAGA API
(Fig.~\ref{fig:arch-overview}, dash arrow), as done by PanDA Broker to submit
MPI scripts. Once scheduled, the pilot's Agent is bootstrapped on Titan's MOM
node and \jhanote{the Agent's Executors on the} worker nodes, and  the
UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
overview}, solid arrow). The Agent's Scheduler schedules the units on one or
more Agent's Executor for execution. The Agent's executors can manage one or
more worker nodes, depending on performance evaluations.

The UnitManager and the Agent communicate via a database that is instantiated on
one of Titan's DTN so as to be reachable by both modules. A similar approach would
be used to enable PanDA Broker to schedule pilots instead of MPI
scripts.\mtnote{Rewrite this after checking it with Andre.}\jhanote{In the interest of space and clarity, I would  remove the last sentence of this para. And merge
the first sentence with the para above.}

@@

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for
communication and coordination of the execution of units. ORTE is a critical
component of the OpenMPI implementation, developed to support distributed
high-performance computing applications operating in a heterogeneous
environment~\cite{XXX}. The system \jhanote{which system? ORTE or OpenMPI in
general?} transparently provides support for interprocess communication,
resource discovery and allocation, and process launching across a variety of
platforms. ORTE provides a mechanism similar to the Pilot concept - it allows
the user to create a \emph{``dynamic virtual machine''} (DVM) that spans
multiple nodes. ORTE provides libraries to enable the submission, monitoring
and managing of tasks, avoiding filesystem bottlenecks and race conditions
with network sockets. As a consequence, it is able to minimize the system
overhead while submitting tasks.

>>>>

\jhanote{not sure MPI script is optimal description or ambiguous.}\mtnote{Is
what we used in the previous section(s) to designate what PanDA Broker submits
to Titan's Torque batch system. Job is already used for what is pulled from
PanDA server.} Multiple generations would also relax two assumptions of the
current execution model: knowing the number of simulations before submitting the
MPI script, and having a fixed number of events per simulation (100 at the
moment). Pilots would enable the scheduling of simulations independently from
whether they were available at the moment of submitting the pilot. Further,
simulations with a varying number of events could be scheduled on a pilot,
depending on the amount of remaining walltime and the distribution of execution
time per event, as shown in \S\ref{ssec:panda_titan},
Figure~\ref{fig:100event-distrib}. These capabilities would increase the
efficiency of the PanDA Broker when there is a large difference between the
number of cores and walltime.

@@

\jhanote{Alternative: "NGE is a runtime system to execute heterogeneous and
dynamically determined tasks that constitute workloads." This definition put the
focus on tasks and not workloads, which is a level above in the RADICAL
conceptual layer diagram.}\mtnote{Done.} NGE is a runtime system to execute
heterogeneous and dynamically determined tasks that constitute workloads.
Fig.~\ref{fig:arch- overview} illustrates its current architecture as deployed
on Titan: the two management modules (Pilot and Unit) represent a simplified
version of the PanDA Broker while the agent module is the pilot submitted to
Titan and executed on its worker nodes. The communication between PanDA Broker
and PanDA Server is abstracted away as it is not immediately useful to evaluate
the performance and capabilities of a pilot on Titan.

@@

NGE exposes an API for the application  \jhanote{workload management?} layer to
describe workloads (Fig.~\ref{fig:arch-overview}, green squares) and pilots
(Fig.~\ref{fig:arch-overview}, red circles), and to instantiate a PilotManager
and a UnitManager.   The PilotManager submits pilots to Titan's Torque
\jhanote{for consistency should we say Moab?}\mtnote{Moab = Scheduler; Torque =
Batch system.} batch system via SAGA API (Fig.~\ref{fig:arch-overview}, dash
arrow), as done by PanDA Broker to submit MPI scripts. Once scheduled, the
pilot's Agent is bootstrapped on Titan's MOM node and \jhanote{the Agent's
Executors on the} the Agent's Executors on the worker nodes, and the
UnitManager schedules units to the Agent's Scheduler (Fig.~\ref{fig:arch-
overview}, solid arrow). The Agent's Scheduler schedules the units on one or
more Agent's Executor for execution. The Agent's executors can manage one or
more worker nodes, depending on performance evaluations. The UnitManager and the
Agent communicate via a database that is instantiated on one of Titan's DTN so
as to be reachable by both modules.

% A similar approach would be used to enable PanDA Broker to schedule pilots
% instead of MPI scripts.\mtnote{Rewrite this after checking it with
% Andre.}
\jhanote{In the interest of space and clarity, I would  remove the last
sentence of this para. And merge the first sentence with the para
above.}\mtnote{Done.}

@@

The NGE Agent uses the \emph{Open Run-Time Environment (ORTE)} for communication
and coordination of the execution of units. ORTE is a critical component of the
OpenMPI implementation, developed to support distributed high-performance
computing applications operating in a heterogeneous environment~\cite{XXX}. The
system \jhanote{which system? ORTE or OpenMPI in general?}\mtnote{I am not sure
but I think ORTE?} transparently provides support for interprocess
communication, resource discovery and allocation, and process launching across a
variety of platforms. ORTE provides a mechanism similar to the Pilot concept -
it allows the user to create a \emph{``dynamic virtual machine''} (DVM) that
spans multiple nodes. ORTE provides libraries to enable the submission,
monitoring and managing of tasks, avoiding filesystem bottlenecks and race
conditions with network sockets. As a consequence, it is able to minimize the
system overhead while submitting tasks.
